---
title: "GLM"
author: "Quinn White"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    df_print: paged
    code_folding: hide
bibliography: "`r rbbt::bbt_write_bib('bibliography_multilevel_report.json', overwrite = TRUE)`"
---



```{css, echo = FALSE}
  #content{
        max-width:2200px;
    }
 
```



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
library(tidyverse)
library(glmnet)
library(here)
```


# Data Cleaning

We used data from the questions `questions.csv` and `clients.csv` for this analysis.

```{r data cleaning}


# source(here('data_cleaning', 'data_cleaning.R'))

questions <- read_csv(here('data', 'questions.csv'))

questions <- questions %>%
  mutate(closed_status = case_when(
    TakenByAttorneyUno != "NULL" & ClosedByAttorneyUno != "NULL"  ~ "Taken and Closed",
    TakenByAttorneyUno == "NULL" ~ "Never Taken",
    TakenByAttorneyUno != "NULL" & 
      ClosedByAttorneyUno == "NULL" ~ 
      "Taken but Not Closed",
    ClosedByAttorneyUno != "NULL" & 
      TakenByAttorneyUno == "NULL" ~ "Closed but Not Taken"
  ),
  success = case_when(closed_status == "Taken and Closed" ~ "Success",
                      closed_status == "Never Taken" ~ "Unsuccessful",
                      closed_status %in% c( "Taken but Not Closed",
                                            "Closed but Not Taken" ) ~ "Unsure")) %>% 
  mutate(across(c(AskedOnUtc,
                  TakenOnUtc, 
                  ClosedOnUtc, 
                  LegalDeadline),
                ~strptime(.,
                          format = "%Y-%m-%d %H:%M:%S", 
                          tz = "GMT") %>%
                  as.POSIXct()))%>%
  mutate(
    asked_on_to_taken_on =
      difftime(TakenOnUtc,AskedOnUtc,
               units = "days"),
    taken_on_to_closed =
      difftime(ClosedOnUtc,TakenOnUtc,
               units = "days")) %>%
  mutate(across(c(asked_on_to_taken_on,taken_on_to_closed),
                as.numeric)) 

#######################################################################################
# THIS SECTION OF CODE IS FROM JOE BRIEN's POST ON STACK OVERFLOW 
# https://stackoverflow.com/questions/9500114/find-which-season-a-particular-date-belongs-to
#######################################################################################
get_season <- function(dates) {
    ws <- as.Date("2020-12-15", format = "%Y-%m-%d") 
    se <- as.Date("2020-3-15",  format = "%Y-%m-%d")
    ss <- as.Date("2020-6-15",  format = "%Y-%m-%d") 
    fe <- as.Date("2020-9-15",  format = "%Y-%m-%d") 

    # Convert dates from any year to 2012 dates
    d <- as.Date(strftime(dates, format="2020-%m-%d"))

    ifelse (d >= ws | d < se, "Winter",
      ifelse (d >= se & d < ss, "Spring",
        ifelse (d >= ss & d < fe, "Summer", "Fall")))
}


questions <- questions %>%
  mutate(season_asked = get_season(AskedOnUtc)) 


# bar plot of Proportion Successful by Season Question was Asked
questions %>%
  group_by(season_asked, success) %>%
  summarize(n=n()) %>%
  group_by(season_asked) %>%
  mutate(tot = sum(n)) %>%
  filter(success=="Success") %>%
  mutate(prop = n/tot) %>%
  ggplot(aes(x=season_asked,y=prop)) +
  geom_bar(stat="identity") +
  theme_bw() +
  labs(title = "Proportion Successful by Season Question was Asked",
       x= "Season",
       y= "Proportion") +
  theme(plot.title = element_text(face="bold", hjust =.5))





clients <- read_csv(here("data", "clients.csv")) %>%
  select(ClientUno, 
         AnnualIncome,MaritalStatus,
         Gender,Age,EthnicIdentity,
         Imprisoned,
         CheckingBalance,
         SavingsBalance,
         InvestmentsBalance) %>%
  mutate(across(c(AnnualIncome, Age), 
                as.numeric))

questions_clients <- questions %>% 
  left_join(clients, by = c('AskedByClientUno' = 'ClientUno')) %>%
  mutate(EthnicIdentity = ifelse(EthnicIdentity == "NULL",
                                 NA, EthnicIdentity)) 



questions_clients <- questions_clients %>%
  filter(AnnualIncome >= 0 & 
           AnnualIncome <= quantile(AnnualIncome,.99,na.rm=TRUE)) %>%
  filter(Age <= 122) %>%
  filter(success != "Unsure")

n_ambiguous <- questions_clients %>% 
  filter(closed_status != "Success" & closed_status != "Unsuccessful") %>%
  nrow()


print(paste0("Proportion of Observations: ", n_ambiguous/nrow(questions_clients) )) 



# pick most frequent gender as reference category
questions_clients %>% 
  group_by(Gender) %>% 
  summarize(n=n()) %>% 
  arrange(desc(n))


questions_clients <- questions_clients %>%
  mutate(across(where(is.character), ~ifelse(. == "NULL", NA, .))) %>%
  mutate(across(c(SavingsBalance, CheckingBalance), as.numeric)) %>%
  select(AnnualIncome, MaritalStatus, 
         Age, Gender, Imprisoned,
         Category, 
         StateAbbr, 
         success, 
         season_asked,
         AskedByClientUno,
         EthnicIdentity) %>%
  mutate(EthnicIdentity = gsub(
    ", Not Hispanic or Latino|Not Hispanic or Latino", '', 
    EthnicIdentity)) %>%
  mutate(EthnicIdentity = case_when(
    EthnicIdentity == "NULL" | EthnicIdentity == "" ~ NA,
    grepl(",", EthnicIdentity) ~ "Multiple Ethnicities Checked",
    TRUE ~ EthnicIdentity
  )) %>%
  mutate(
         Gender = case_when(
           Gender == "Female" ~ "Female",
            Gender == "Male" ~ "Male",
          grepl("Non-Conforming", Gender,fixed=TRUE) ~ "Non-Conforming",
          TRUE ~ "Other"),
         success= case_when(
           success == "Unsuccessful" ~ 0, 
           success=="Success" ~ 1)) %>%
  mutate(across(c(MaritalStatus,
                  Gender,
                  Imprisoned,
                  success,
                  Category,
                  StateAbbr, EthnicIdentity, season_asked), as.factor)) %>%
  mutate(MaritalStatus = relevel(MaritalStatus,
                                 ref = "Single"),
         Gender = relevel(Gender, 
                          ref = "Female"),
         Imprisoned = relevel(Imprisoned,
                              ref= "No"),
         EthnicIdentity = relevel(EthnicIdentity, 
                                  ref="Caucasian"),
         Category = relevel(Category,
                            ref = "Family and Children"),
          StateAbbr = relevel(StateAbbr, 
                              ref = "TX"),
          Category = relevel(Category, ref= "Family and Children"),
         season_asked = relevel(season_asked, 
                                ref = "Summer")) %>%
  mutate(across(where(is.numeric),
                ~ ifelse(is.na(.), mean(., na.rm = TRUE), .))) %>%
  filter(if_all(.cols = everything(),  ~ !is.na(.x)))




num_more_than_one <- questions_clients %>% 
  group_by(AskedByClientUno) %>%
  mutate(n=n()) %>% 
  filter(n>1) %>%
  nrow()



prop <- num_more_than_one / nrow(questions_clients)

cat(paste("Number of Observations Corresponding to Client ID that Asked More than 1: ", prop))

num_more_than_2 <- questions_clients %>% 
  group_by(AskedByClientUno) %>% 
  summarize(n=n()) %>% 
  filter(n>2) %>% 
  nrow()

prop_2 <- num_more_than_2 / nrow(questions_clients)

cat(paste("Number of Observations Corresponding to Client ID that Asked More than 2: ", prop_2))

train <- questions_clients %>%
  select(-c(AskedByClientUno)) 


# questions_clients$MaritalStatus %>% unique()

```


# Formulation of the Model

We fit a logistic regression model with:

* Annual Income (standardized)
* Marital Status 
* Age (standardized)
* Gender
* Imprisoned
* Category
* State Abbreviation
* Ethnic Identity 
* Season Question was Asked

Choice of standardization following the approach of @gelman2008.


```{r run logistic regression}

# standardize continuous predictors
train <- train %>% 
  mutate(AnnualIncome_std = (AnnualIncome - mean(AnnualIncome))/ (2*sd(AnnualIncome)),
         Age_std = (Age -  mean(Age)/ (2*sd(Age)))
         ) 

# mod <- glm( success ~ AnnualIncome + MaritalStatus + 
#        Age + Gender + Imprisoned  +
#        Category + StateAbbr, data = train, family = "binomial")

mod <- glm( success ~ AnnualIncome_std + MaritalStatus + 
       Age_std + Gender + Imprisoned  +
       Category + StateAbbr + EthnicIdentity + season_asked,
       data = train, family = "binomial")


train %>% 
  group_by(Category, success) %>%
  summarize(n=n()) %>%
  mutate(success=ifelse(success == 1, "Success", "Unsuccessful")) %>%
  group_by(Category) %>%
  mutate(tot =sum(n)) %>%
  mutate(prop = n/tot) %>%
  filter(success=="Success") %>%
  ggplot(aes(x= fct_reorder(Category,prop), y = prop)) +
  geom_bar(stat="identity", position ="dodge") +
  coord_flip()

train %>%
  group_by(Category) %>%
  summarize(n=n())




train %>% 
  mutate(predicted =predict.glm(mod, train, 
                                type = "response")) %>%
  mutate(closed_pred = ifelse(predicted > 0.5, 1, 0),
         error = case_when(
           success == 0 & closed_pred == 0 ~ "True negative",
           success == 0 & closed_pred == 1 ~ "False Positive",
          success == 1 & closed_pred == 1 ~ "True Positive",
          success == 1 & closed_pred == 0 ~ "False Negative"
         )) %>%
  group_by(error) %>% 
  summarize(n=n()) %>%
  ungroup() %>%
  mutate(tot = sum(n)) %>%
  mutate(n=n/tot)


```

```{r}
library(pROC)

test_prob <- predict(mod, newdata = train, type = "response")
test_roc <- roc(train$success ~ test_prob, plot = TRUE, print.auc = TRUE)

```

```{r, error = TRUE, fig.width= 18, fig.height = 16}
  
# mod$coefficients %>%
#   as_tibble(rownames = "Variable") %>%
#   filter(Variable != "(Intercept)") 
#  
#   
sig <- summary(mod)$coefficients %>% 
  as_tibble(rownames="Variable") %>%
   filter(`Pr(>|z|)` < 0.05 & Variable != "(Intercept)") %>%
  pull(Variable)


# conf <- confint(mod)
# saveRDS(conf, here("data/confint.RDS"))


# 
# summary(mod)$coefficients %>%
#   as_tibble(rownames= "Variable") %>%
#   filter(Variable != "(Intercept)") 


coefs <- mod$coefficients %>%
  as_tibble(rownames = "Variable") %>%
  mutate(value =exp(value)) %>%
  select(coefficient = value, Variable)


mod$coefficients %>%
  as_tibble(rownames = "Variable") %>%
  mutate(Variable = gsub("StateAbbr", "", Variable)) %>%
  mutate(state = ifelse(nchar(Variable) != 2,
                        "Coefficients",
                        "State Coefficients")) %>%
  filter(Variable != "(Intercept)")  %>%
  ggplot(aes(y = fct_reorder(Variable, value),
             x= value)) +
  geom_point() +
  theme_bw() +
  facet_wrap(~state, scales="free") +
  geom_vline(xintercept = 0,
             linetype = 2, 
             color="darkred",
             alpha = .5) +
  labs(y = "")



# confidence intervals
ci <- confint.default(mod, method = "wald") %>%
    as_tibble(rownames = "Variable") %>%
    left_join(coefs) %>%
  mutate(across(c( `2.5 %`,`97.5 %`), exp)) 


saveRDS(ci, here('results', "confint_model.RDS"))
  
  
ci %>%
    filter(Variable %in% sig) %>%
    mutate(Variable = gsub("StateAbbr", "", Variable)) %>%
    mutate(coefficient_type = case_when(
      nchar(Variable) == 2 ~ "State Coefficients",
     #  grepl("Gender", Variable) ~ "Gender",
      grepl("Category", Variable) ~ "Question Category",
      grepl("EthnicIdentity", Variable) ~ "Ethnic Identity",
     grepl("season_asked", Variable) ~"Season",
   #   grepl("Marital", Variable) ~ "Marital",
      TRUE ~ "Other Demographic Characteristics")) %>%
  mutate(Variable = gsub("Age_std", "Age (Standardized)",
                         Variable),
         Variable = gsub("MaritalStatus", 
                         "Marital Status: ",
                         Variable),
         Variable = gsub("Gender|EthnicIdentity|season_asked", '', Variable),
         Variable = gsub('AnnualIncome_std',
                         'Annual Income (Standardized)',
                         Variable),
         Variable = gsub('Category',
                         '', Variable)) %>%
    filter(Variable != "(Intercept)") %>%
  ggplot(aes(y = fct_reorder(Variable, `2.5 %`),
             xmin = `2.5 %`,
             xmax =`97.5 %`)) +
  geom_point(color = "darkred", aes(x=coefficient)) +
  geom_errorbar(width = .5) +
  theme_bw() +
  facet_wrap(~coefficient_type, 
             scales="free", 
             ncol =2) +
  geom_vline(xintercept = 1,
             linetype = 2, 
             color="darkred",
             alpha = .5) +
  labs(y = "",
       x = "Odds Ratio",
       title = "Coefficient Estimates for Logistic Regression Model") +
  theme(strip.background = element_rect(fill = "#323232"),
        strip.text=element_text(color="white",
                                face="bold",
                                size = 14),
        axis.text = element_text(size = 12),
        plot.title = element_text(hjust = .5, face = "bold", size = 25))




```

```{r, results='asis', eval=FALSE}


l <- ceiling(length(mod$coefficients)/2)
end <- length(mod$coefficients) -1

model_for1 <- paste0("\\beta_{", 0:l, "} (",
                gsub("_", ".", names(mod$coefficients)[1:(l+1)], fixed = TRUE ),
                ")",
                collapse="+") %>%
  gsub("((Intercept))", "",., fixed=TRUE)

model_for2 <-  paste0("\\beta_{", (l+1):end, "} (",
                gsub("_", ".", names(mod$coefficients)[(l+2):end], 
                     fixed = TRUE ),
                ")",
                collapse="+") 

cat("$$\\hat{\\text{log odds}(y=success)}=", 
    model_for1, "+$$")
cat("$$", model_for2, "$$")


```


# Model Diagnostics


For model diagnostics, we calculate the randomized quantile residuals using the implementationof the [DHARMa R package](https://cran.r-project.org/web/packages/DHARMa/vignettes/DHARMa.html) [@florianhartig2022]. Randomized quantile residuals were introduced by @dunn1996a and offer advantages over other forms of residuals (e.g., deviance and Pearson) for diagnosing problems with generalized linear models.


```{r, fig.show='hold', out.width='50%'}

# train %>% group_by(Category) %>%
#   summarize(n=n()) %>% 
#   arrange(desc(n))

library(DHARMa)


# pearson residuals
train %>%
  mutate(res = residuals(mod, type = "pearson"),
         odds = predict(mod, type = "link"),
        ) %>%
  ggplot(aes(sample=res)) +
  stat_qq() +
  stat_qq_line(color = "darkred") +
  labs(title = "Pearson Residuals")+
  theme_bw() +
  theme(plot.title = element_text(hjust = .5, 
                                  face = "bold",
                                  size = 16))



# randomized quantile residuals
simulationOutput <- simulateResiduals(fittedModel = mod,
                                      plot = FALSE, 
                                      n = 250)

tibble(residuals = residuals(simulationOutput)) %>%
  ggplot(aes(sample = residuals)) +
  stat_qq(distribution = stats::qunif, alpha = .01, size = 1) +
  stat_qq_line(distribution = stats::qunif,
               color = "darkred",
               linewidth = 1.2) +
  theme_bw() +
  labs(title = "QQ Plot of Randomized Quantile Residuals") +
  theme(plot.title = element_text(hjust = .5, 
                                  face = "bold",
                                  size = 16))



```


# Frequency of Different Variables in the Data


```{r, fig.show='hold', out.width='50%'}

library(scales)



questions_clients %>% 
  group_by(EthnicIdentity) %>% 
  summarize(n=n()) %>%
  ggplot(aes(x= fct_reorder(EthnicIdentity,n),
             y=n)) +
  geom_bar(stat="identity") +
  theme_bw() +
  labs(x = "Number of Observations",
       y = "Ethnic Identity",
       title = "Number of Observations by Ethic Identity") +
  coord_flip() +
  scale_y_continuous(labels=comma) +
  theme(plot.title = element_text(hjust = .5, 
                                  face = "bold",
                                  size = 16))
  
  
questions_clients %>% 
  group_by(season_asked) %>% 
  summarize(n=n()) %>%
  ggplot(aes(x= fct_reorder(season_asked,n),
             y=n)) +
  geom_bar(stat="identity") +
  theme_bw() +
  labs(x = "Number of Observations",
       y = "Season Asked",
       title = "Number of Observations by Season Asked") +
  coord_flip() +
  scale_y_continuous(label=comma)+
  theme(plot.title = element_text(hjust = .5, 
                                  face = "bold",
                                  size = 16))
  


questions_clients %>% 
  group_by(Category) %>% 
  summarize(n=n()) %>%
  ggplot(aes(x= fct_reorder(Category,n),
             y=n)) +
  geom_bar(stat="identity") +
  theme_bw() +
  labs(x = "Number of Observations",
       y = "Question Category",
       title = "Number of Observations by Question Category") +
  coord_flip() +
  scale_y_continuous(label=comma)+
  theme(plot.title = element_text(hjust = .5, 
                                  face = "bold",
                                  size = 16))


```

# References

